Classical density functional theory for the square-well liquid

### 2014-05-08 papers/square-well-liquid/paper.tex
switch to using even fewer spheres in quick computation Tested on bennet. Took 0:17.92.
started TPT section for SW paper Tested on elliot. Took 1:14.07.

\author{Student One}
\author{Michael MI? Perlin}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}

@@

@@

\fixme{Add here a complete exposition of this functional form,
  reproducing what Gil Villegas says.}

\fixme{DRoth -- Some notes are as follows:}

@@

Then, the high-temp expansion is an expansion of the monomer term:
\begin{align}
  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \cdots
\end{align}
\fixme{I'm not sure how to express this in terms of $F_\text{disp}[n]$ alone.}

The term $a_1$ is given by
\begin{align}
  a_1 &= -2\pi\rho\epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
  &= -3\rho b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
\end{align}
If assume a random correlations between the particles' positions, for all distances, we have $g^\text{HS}(r) = 1$. This yields
\begin{align}
  a_1^\text{VDW} = -\rho\alpha^\text{VDW},
\end{align}
the van der Waals mean-field energy.

The second-order term is quite a bit more complicated; one must have
knowledge of higher-order correlation functions to determine
$a_2$. This term describes the response of the attractive energy due
to the compression of the fluid from the attractive well. Gil-Villegas
\emph{et al.} base their expression on an approximation by Barker and
Henderson\cite{barker1967-SW-perturbation}. \fixme{Gil-Villegas also
  cites 2 other Barker and Henderson papers, but I can't find one and
  the other doesn't seem relevant}. Barker and Henderson's
approximation considers fluctuations in the number of particles in the
potential well. In this way, the fluctuations in $a_2$ are related to the pressure and compressibility
of the liquid. Given pressure $P^\text{HS}$ and isothermal compressibility $K^\text{HS} = kT\left(\partial\rho /\partial P^\text{HS}\right)_T$, we have:
\begin{align}
  a_2 &= -\pi\rho\epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial\rho g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
  &= -\pi\rho\epsilon^2K^\text{HS}\frac{\partial}{\partial\rho}\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2\rho g^\text{HS}\,dr\right] \\
  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial\rho}\left[-3\rho b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
\end{align}

@@

\fixme{End of my comments --DRoth}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/periodic-ww13-ff30-N200-E.pdf}
  \caption{A simple DOS plot!}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/periodic-ww30-ff30-N200-E.pdf}
  \caption{A simple DOS plot!}
\end{figure}

>>>>

\author{Student One}
\author{Michael A. Perlin}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}

@@

\fixme{And introduction to TPT---why do we use it?}

\fixme{For now, the discussion follows entirely from the Simple Liquids book~\cite{simple-liquids}}

The pair potential $V(\rr)$ describes the interaction between two hard spheres. For the square-well model,
\begin{align}
  v_\text{SW}(\rr) &=
    \begin{cases}
      \infty & r < \sigma \\
      -\epsilon & \sigma < r < \lambda\sigma \\
      0 & \lambda\sigma < r
    \end{cases}
\end{align}

Now, consider this to be the reference, and add a small perturbation:
\begin{align}
  v_\lambda(\rr) &= v_\text{SW}(\rr) + \lambda w(\rr)
\end{align}

The total potential energy of our system is
\begin{align}
  V_N(\lambda) = \sum_\rr v_\lambda(\rr)
\end{align}

@@

\fixme{The following hopes to be a complete exposition of this functional form,
  reproducing what Gil Villegas says.}

@@

Then, the high-temp expansion is an expansion of the monomer term:
\begin{align}
  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \cdots
\end{align}

The term $a_1$ is given by
\begin{align}
  a_1 &= -2\pi n \epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
  &= -3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
\end{align}
If we assume random correlations between the particles' positions, for all distances, we have $g^\text{HS}(r) = 1$. This yields
\begin{align}
  a_1^\text{VDW} = - n \alpha^\text{VDW},
\end{align}
the van der Waals mean-field energy.

The second-order term is quite a bit more complicated; one must have
knowledge of higher-order correlation functions to determine
$a_2$. This term describes the response of the attractive energy due
to the compression of the fluid from the attractive well. Gil-Villegas
\emph{et al.} base their expression on an approximation by Barker and
Henderson\cite{barker1967-SW-perturbation}. Barker and Henderson's
approximation considers fluctuations in the number of particles in the
potential well. In this way, the fluctuations in $a_2$ are related to
the pressure and compressibility of the liquid. Given pressure
$P^\text{HS}$ and isothermal compressibility $K^\text{HS} =
kT\left(\partial n  /\partial P^\text{HS}\right)_T$, we have:
\begin{align}
  a_2 &= -\pi n \epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
  &= -\pi n \epsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \\
  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
\end{align}

@@

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E.pdf}
  \caption{A simple DOS plot!}
\end{figure}

%% \begin{figure}
%%   \in cludegraphics[width=\columnwidth]{figs/periodic-ww300-ff30-N200-E.pdf}
%%   \caption{A simple DOS plot!}
%% \end{figure}
